---
title: "TL_Trans Detailed Analysis"
author: "Taylor Lindsay"
date: "2023-08-21"
output:
  pdf_document: default
  word_document: default
  html_document: default
---
\newpage
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = FALSE)
```
# Summary & Terms 
This document shows the results from a reciprocal transplant experiment in Puerto Rico 

OFAV - *Orbicella faveolata*, OFRA - *Orbicella franksii* 

PP - Deep to deep transplant, PS - Deep to shallow transplant, SS - Shallow to shallow transplant, SP - Shallow to deep transplant 

# Load libraries: Done  
```{r libraries & imports, include=FALSE}
# Libraries 
library(tidyverse)
library(dplyr)
library(ggpubr)
library(ggplot2)

# load data & clean it up 

raw <- read.csv('~/Desktop/GITHUB/TLPR21/TL_Trans_Results.csv')

# make a 'group column' that allows us to facet wrap 
raw <- raw %>%
  mutate(group = str_sub(full_treatment,1,6))

# make a table of means & SD 
means <- raw %>%
  group_by(full_treatment) %>%
  summarise(across(everything(), .f = list(mean = mean, sd = sd), na.rm = TRUE))
  
# Define my groups for stat compare means 
treatment_comparisons <- list( c("OFAV_PS","OFAV_PP"), c("OFAV_SP","OFAV_SS"), c("OFRA_PP","OFRA_PS"), c("OFRA_PP","OFAV_PP"), c("OFAV_PP","OFAV_SS"))

# set orders for graphs 
measurement_order <- c('OFAV_S','OFAV_P','OFRA_P') 
x_order <- c('OFAV_SP','OFAV_SS','OFAV_PP', 'OFAV_PS', 'OFRA_PP', 'OFRA_PS') 

# define offset of points 
pj <- position_jitterdodge(jitter.width=0.1, seed=9,
                           jitter.height = 0)
```
# Remove outliers: Done
Each variable in this data set has been filtered for outliers > 3x z-score 

# Check for Normalcy: Done
Run Shapiro and Bartlett tests on all parameters to test for normality and homogeneity 

**Results of Normalcy check**
All physiology metrics do not have parametric shapes or homogeneity, must use non-parametric tests on all analyses 

t-test -> Wilcoxon's 


\newpage
**Chlorophyll a**
```{r chla_normalcy, fig.height = 2, fig.width = 2, fig.align = "center", fig.show="hold"}
ggdensity(raw$chla.ug.cm2, xlab = "chla.ug.cm2")
ggqqplot(raw$chla.ug.cm2)
shapiro.test(raw$chla.ug.cm2)
bartlett.test(chla.ug.cm2~full_treatment, data=raw) 
```
**Chlorophyll c2**
```{r chlc2_normalcy, fig.height = 2, fig.width = 2, fig.align = "center", fig.show="hold"}
ggdensity(raw$chlc2.ug.cm2, xlab = "chlc2.ug.cm2")
ggqqplot(raw$chlc2.ug.cm2)
shapiro.test(raw$chlc2.ug.cm2)
bartlett.test(chlc2.ug.cm2~full_treatment, data=raw) 
```
\newpage
**Symbiont Density**
```{r sym_normalcy, fig.height = 2, fig.width = 2, fig.align = "center", fig.show="hold"}
ggdensity(raw$sym.cm2, xlab = "sym.cm2")
ggqqplot(raw$sym.cm2)
shapiro.test(raw$sym.cm2)
bartlett.test(sym.cm2~full_treatment, data=raw) 
```
**Protein Concentration** 
```{r prot_normalcy, fig.height = 2, fig.width = 2, fig.align = "center", fig.show="hold"}
ggdensity(raw$prot_mg.cm2, xlab = "prot_mg.cm2")
ggqqplot(raw$prot_mg.cm2)
shapiro.test(raw$prot_mg.cm2)
bartlett.test(prot_mg.cm2~full_treatment, data=raw) 
```
\newpage
**Symbiont Ash Free Dry Weight**
```{r afdw_sym_normalcy, fig.height = 2, fig.width = 2, fig.align = "center", fig.show="hold"}
ggdensity(raw$Sym_AFDW_mg.cm2, xlab = "Sym_AFDW_mg.cm2")
ggqqplot(raw$Sym_AFDW_mg.cm2)
shapiro.test(raw$Sym_AFDW_mg.cm2)
bartlett.test(Sym_AFDW_mg.cm2~full_treatment, data=raw) 
```
**Host Ash Free Dry Weight** 
```{r afdw_host_normalcy, fig.height = 2, fig.width = 2, fig.align = "center", fig.show="hold"}
ggdensity(raw$Host_AFDW_mg.cm2, xlab = "Host_AFDW_mg.cm2")
ggqqplot(raw$Host_AFDW_mg.cm2)
shapiro.test(raw$Host_AFDW_mg.cm2)
bartlett.test(Host_AFDW_mg.cm2~full_treatment, data=raw) 
```


\newpage
# Chlorophyl a
```{r chla, fig.height = 3, fig.width = 6, fig.align = "center"}
chla <- ggplot(raw, aes(x=factor(full_treatment, level=x_order), y=chla.ug.cm2, fill=factor(group, level=measurement_order))) +
  geom_boxplot(alpha=0.5, outlier.size=0) +
  geom_point(aes(color=factor(group, level=measurement_order)), position = pj, size=1, show.legend = FALSE)+
  scale_fill_manual(values = c("#EBB134", "#ED6B5F", "#6060A8"), labels=c('OFAV Shallow', 'OFAV Deep', 'OFRA Deep')) + 
  scale_color_manual(values = c("#EBB134", "#ED6B5F", "#6060A8")) + 
  labs(y= "Chlorophyl a (ug/cm2)", x = "Depth Treatment", fill='Origin') + 
  scale_x_discrete(labels=c('Deep', 'Shallow', 'Deep', 'Shallow', 'Deep', 'Shallow')) + 
  theme_classic() +
  stat_compare_means(comparisons = treatment_comparisons, method = "wilcox.test", 
                     symnum.args = list(cutpoints = c(0, 0.001 ,0.01, 0.05, Inf), symbols = c("***", "**", "*", "")))
chla
pairwise.wilcox.test(raw$chla.ug.cm2, raw$full_treatment, p.adjust.method="none")
```

### Results summary 

**Expected trend: low in shallow** 

There were no differences in chl a between species controls or deep and shallow controls of OFAV. *this is surprising* 

For both OFAV and OFRA, there was a significant decrease in chl a concentration when samples were transplanted from deep to shallow. There was no effect of transplanting OFAV from shallow to deep. *this follows our expected trend, with the exception that OFAV transplanted to deep expressed no change* 

\newpage
# Chlorophyll c2

```{r chlc, fig.height = 3, fig.width = 6, fig.align = "center"}
chlc <- ggplot(raw, aes(x=factor(full_treatment, level=x_order), y=chlc2.ug.cm2, fill=factor(group, level=measurement_order))) +
  geom_boxplot(alpha=0.5, outlier.size=0) +
  geom_point(aes(color=factor(group, level=measurement_order)), position = pj, size=1, show.legend = FALSE)+
  scale_fill_manual(values = c("#EBB134", "#ED6B5F", "#6060A8"), labels=c('OFAV Shallow', 'OFAV Deep', 'OFRA Deep')) + 
  scale_color_manual(values = c("#EBB134", "#ED6B5F", "#6060A8")) + 
  labs(y= "Chlorophyl c2 (ug/cm2)", x = "Depth Treatment", fill='Origin') + 
  scale_x_discrete(labels=c('Deep', 'Shallow', 'Deep', 'Shallow', 'Deep', 'Shallow')) + 
  theme_classic() +
  stat_compare_means(comparisons = treatment_comparisons, method = "wilcox.test", 
                     symnum.args = list(cutpoints = c(0, 0.001 ,0.01, 0.05, Inf), symbols = c("***", "**", "*", "")))
chlc
pairwise.wilcox.test(raw$chlc2.ug.cm2, raw$full_treatment, p.adjust.method="none")
```

### Results summary 

**Expected results: low in shallow** 

There were no differences in chl c2 between species controls or deep and shallow controls of OFAV. *surprising, we would expect depth related changes* 

OFRA experienced a significant decrease in chl c2 when transplanted from deep to shallow.  *this follows the expected trend*

OFAV experienced a significant decrease in chl c2 when transplanted from shallow to deep. *this is the opposite of the expected trend* 

*it doesn't seem like c2 follows the same trends that chl a follows* 

\newpage
# Symbiont density 
```{r sym,  fig.height = 3, fig.width = 6, fig.align = "center"}
sym <- ggplot(raw, aes(x=factor(full_treatment, level=x_order), y=sym.cm2, fill=factor(group, level=measurement_order))) +
  geom_boxplot(alpha=0.5, outlier.size=0) +
  geom_point(aes(color=factor(group, level=measurement_order)), position = pj, size=1, show.legend = FALSE)+
  scale_fill_manual(values = c("#EBB134", "#ED6B5F", "#6060A8"), labels=c('OFAV Shallow', 'OFAV Deep', 'OFRA Deep')) + 
  scale_color_manual(values = c("#EBB134", "#ED6B5F", "#6060A8")) + 
  labs(y= "Symbiont Density (count/cm2)", x = "Depth Treatment", fill='Origin') + 
  scale_x_discrete(labels=c('Deep', 'Shallow', 'Deep', 'Shallow', 'Deep', 'Shallow')) + 
  theme_classic() +
  stat_compare_means(comparisons = treatment_comparisons, method = "wilcox.test", 
                     symnum.args = list(cutpoints = c(0, 0.001 ,0.01, 0.05, Inf), symbols = c("***", "**", "*", "")))
sym
pairwise.wilcox.test(raw$sym.cm2, raw$full_treatment, p.adjust.method="none")
```

### Results summary 

**Expected results: studies have found mixed results, but most suggest that there are fewer symbionts in shallow environments**

There were no differences in symbiont density between species controls or deep and shallow controls of OFAV. 

For both OFAV and OFRA that were transplanted from deep to shallow, there was no effect on the symbiont density. 

For OFAV transplanted from shallow to deep, there was a significant decrease in symbiont density. *this goes against expected results* 

\newpage
# Protein
```{r protein,  fig.height = 3, fig.width = 6, fig.align = "center"}
protein <- ggplot(raw, aes(x=factor(full_treatment, level=x_order), y=prot_mg.cm2, fill=factor(group, level=measurement_order))) +
  geom_boxplot(alpha=0.5, outlier.size=0) +
  geom_point(aes(color=factor(group, level=measurement_order)), position = pj, size=1, show.legend = FALSE)+
  scale_fill_manual(values = c("#EBB134", "#ED6B5F", "#6060A8"), labels=c('OFAV Shallow', 'OFAV Deep', 'OFRA Deep')) + 
  scale_color_manual(values = c("#EBB134", "#ED6B5F", "#6060A8")) + 
  labs(y= "Protein (mg/cm2)", x = "Depth Treatment", fill='Origin') + 
  scale_x_discrete(labels=c('Deep', 'Shallow', 'Deep', 'Shallow', 'Deep', 'Shallow')) + 
  theme_classic() +
  stat_compare_means(comparisons = treatment_comparisons, method = "wilcox.test", 
                     symnum.args = list(cutpoints = c(0, 0.001 ,0.01, 0.05, Inf), symbols = c("***", "**", "*", "")))
protein
pairwise.wilcox.test(raw$prot_mg.cm2, raw$full_treatment, p.adjust.method="none")
```

### Results summary 

**Expected results: higher protein concentrations in shallow environments**

There was no difference in protein concentration between deep and shallow controls of OFAV. *this is suprising because our results showed a strong effect of depth on protein concentration. It is possible that this change in protein occurs quickly during the period of acclimation, then levels out over time* 

In deep samples only, OFAV controls had a significantly higher concentration of protein than OFRA controls. *this makes sense because OFAV is a generalist that can survive in shallow environments, where we would expect the protein to be higher, while OFRA is a depth-specialist and therefore would usually have lower protein at depth*

All three treatment pairs experienced a significant change in protein concentration between shallow and deep. No matter which direction the transplant occurred, shallow samples always had higher protein concentrations than deep for both OFAV and OFRA. *this follows our expectations that shallow samples have higher levels of protein* 

\newpage
# Symbiont Biomass 
```{r AFDW sym, fig.height = 3, fig.width = 6, fig.align = "center"}
AFDW_sym <- ggplot(raw, aes(x=factor(full_treatment, level=x_order), y=Sym_AFDW_mg.cm2, fill=factor(group, level=measurement_order))) +
  geom_boxplot(alpha=0.5, outlier.size=0) +
  geom_point(aes(color=factor(group, level=measurement_order)), position = pj, size=1, show.legend = FALSE)+
  scale_fill_manual(values = c("#EBB134", "#ED6B5F", "#6060A8"), labels=c('OFAV Shallow', 'OFAV Deep', 'OFRA Deep')) + 
  scale_color_manual(values = c("#EBB134", "#ED6B5F", "#6060A8")) + 
  labs(y= "Symbiont AFDW (mg/cm2)", x = "Depth Treatment", fill='Origin') + 
  scale_x_discrete(labels=c('Deep', 'Shallow', 'Deep', 'Shallow', 'Deep', 'Shallow')) + 
  theme_classic() +
  stat_compare_means(comparisons = treatment_comparisons, method = "wilcox.test", 
                     symnum.args = list(cutpoints = c(0, 0.001 ,0.01, 0.05, Inf), symbols = c("***", "**", "*", "")))
AFDW_sym
pairwise.wilcox.test(raw$Sym_AFDW_mg.cm2, raw$full_treatment, p.adjust.method="none")
```

### Results summary 

**Expected results: dependent on symbiont density, which we expect to be higher at depth** 

OFAV control samples had significantly higher biomass in shallow sites than deep. *goes against expectations*

OFRA and OFAV deep controls did not differ.

For both OFAV and OFRA, there was a significant increase in symbiont biomass when transplanted to shallow environments. *goes against expectations*

OFRA did not experience significant change in biomass when transplanted to shallow. 


# Host Biomass 
```{r AFDW host, fig.height = 3, fig.width = 6, fig.align = "center"}
AFDW_host <- ggplot(raw, aes(x=factor(full_treatment, level=x_order), y=Host_AFDW_mg.cm2, fill=factor(group, level=measurement_order))) +
  geom_boxplot(alpha=0.5, outlier.size=0) +
  geom_point(aes(color=factor(group, level=measurement_order)), position = pj, size=1, show.legend = FALSE)+
  scale_fill_manual(values = c("#EBB134", "#ED6B5F", "#6060A8"), labels=c('OFAV Shallow', 'OFAV Deep', 'OFRA Deep')) + 
  scale_color_manual(values = c("#EBB134", "#ED6B5F", "#6060A8")) + 
  labs(y= "Host AFDW (mg/cm2)", x = "Depth Treatment", fill='Origin') + 
  scale_x_discrete(labels=c('Deep', 'Shallow', 'Deep', 'Shallow', 'Deep', 'Shallow')) + 
  theme_classic() +
  stat_compare_means(comparisons = treatment_comparisons, method = "wilcox.test", 
                     symnum.args = list(cutpoints = c(0, 0.001 ,0.01, 0.05, Inf), symbols = c("***", "**", "*", "")))
AFDW_host
pairwise.wilcox.test(raw$Host_AFDW_mg.cm2, raw$full_treatment, p.adjust.method="none")
```
### Results summary 

**Expected trend: biomass is higher in shallow corals** 

OFAV has higher biomass in their shallow control than their deep control. *this follows our expected trend* 

OFRA has significantly lower biomass than OFAV. *this follows the expected trend with OFRA being a deepth specialist, and OFAV being more generalist*

OFRA transplanted from deep to shallow experienced an increase in biomass, while the same treatment on OFAV had no change. *this follows our expected trend, with OFAV sample size too low but still showing the trend* 

OFAV transplanted from shallow to deep experienced a decrease in biomass. *this follows our expected trend* 

\newpage
# Calice Density  

```{r density, fig.height = 3, fig.width = 6, fig.align = "center"}

# graph calice density 
density <- ggplot(raw, aes(x=factor(full_treatment, level=x_order), y=D, fill=factor(group, level=measurement_order))) +
  geom_boxplot(alpha=0.5, outlier.size=0) +
  geom_point(aes(color=factor(group, level=measurement_order)), position = pj, size=1, show.legend = FALSE)+
  scale_fill_manual(values = c("#EBB134", "#ED6B5F", "#6060A8"), labels=c('OFAV Shallow', 'OFAV Deep', 'OFRA Deep')) + 
  scale_color_manual(values = c("#EBB134", "#ED6B5F", "#6060A8")) + 
  labs(y= "Calice Density (count/cm2)", x = "Depth Treatment", fill='Origin') + 
  scale_x_discrete(labels=c('Deep', 'Shallow', 'Deep', 'Shallow', 'Deep', 'Shallow')) + 
  theme_classic() +
  stat_compare_means(comparisons = treatment_comparisons, method = "wilcox.test", 
                     symnum.args = list(cutpoints = c(0, 0.001 ,0.01, 0.05, Inf), symbols = c("***", "**", "*", "")))
density
pairwise.wilcox.test(raw$D, raw$full_treatment, p.adjust.method="none")

```

### Results summary 

**Expected trend: Calices are smaller in shallow environments, so higher density** 

OFAV controls were significanly higher density in shallow environments *This follows our expected results.*

OFAV deep control was significantly higher density than OFRA deep control *This follows our expected results.*

For both OFAV and OFRA when transplanted from deep to shallow, there was an increase in calice density, while there was no change when transplanted from shallow to deep. *This follows our expected results.* 

\newpage
# Calice Area 
```{r Area, fig.height = 3, fig.width = 6, fig.align = "center"}

# graph calice density 
area <- ggplot(raw, aes(x=factor(full_treatment, level=x_order), y=A, fill=factor(group, level=measurement_order))) +
  geom_boxplot(alpha=0.5, outlier.size=0) +
  geom_point(aes(color=factor(group, level=measurement_order)), position = pj, size=1, show.legend = FALSE)+
  scale_fill_manual(values = c("#EBB134", "#ED6B5F", "#6060A8"), labels=c('OFAV Shallow', 'OFAV Deep', 'OFRA Deep')) + 
  scale_color_manual(values = c("#EBB134", "#ED6B5F", "#6060A8")) + 
  labs(y= "Coralite Area (cm2)", x = "Depth Treatment", fill='Origin') + 
  scale_x_discrete(labels=c('Deep', 'Shallow', 'Deep', 'Shallow', 'Deep', 'Shallow')) + 
  theme_classic() +
  stat_compare_means(comparisons = treatment_comparisons, method = "wilcox.test", 
                     symnum.args = list(cutpoints = c(0, 0.001 ,0.01, 0.05, Inf), symbols = c("***", "**", "*", "")))
Carea <- ggplot(raw, aes(x=factor(full_treatment, level=x_order), y=CA, fill=factor(group, level=measurement_order))) +
  geom_boxplot(alpha=0.5, outlier.size=0) +
  geom_point(aes(color=factor(group, level=measurement_order)), position = pj, size=1, show.legend = FALSE)+
  scale_fill_manual(values = c("#EBB134", "#ED6B5F", "#6060A8"), labels=c('OFAV Shallow', 'OFAV Deep', 'OFRA Deep')) + 
  scale_color_manual(values = c("#EBB134", "#ED6B5F", "#6060A8")) + 
  labs(y= "Calice Area (cm2)", x = "Depth Treatment", fill='Origin') + 
  scale_x_discrete(labels=c('Deep', 'Shallow', 'Deep', 'Shallow', 'Deep', 'Shallow')) + 
  theme_classic() +
  stat_compare_means(comparisons = treatment_comparisons, method = "wilcox.test", 
                     symnum.args = list(cutpoints = c(0, 0.001 ,0.01, 0.05, Inf), symbols = c("***", "**", "*", "")))
area
Carea
pairwise.wilcox.test(raw$A, raw$full_treatment, p.adjust.method="none")
pairwise.wilcox.test(raw$CA, raw$full_treatment, p.adjust.method="none")
```


### Results summary 

**Expected trend: Calices are smaller in shallow environments, but morphology will change very slowly and therefore we won't be surprised by small change** 

For both calice area and corallite area, OFAV controls were significanly smaller in shallow environments *This follows our expected results.*

For both calice area and corallite area, OFAV deep control were significantly smaller than OFRA deep control *This follows our expected results.*

For each of the OFAV treatments, there was no change in corallite or calice area. *This is expected because these species are so slow growing* 

For OFRA, there was no significant change in corallite area when transplanted from deep to shallow, but there was a significant decrease in the calice size. *This is surprising given how slow growing the species are, but follows expected patterns of smaller calices in shallow. We hypothesize that the corals acclimated to a foreign environment very quickly by reducing photoinhibition through reflectance of the calice.* 

\newpage
# Calice diameter 
```{r diameter, fig.height = 3, fig.width = 6, fig.align = "center"}

# graph calice density 
diameter <- ggplot(raw, aes(x=factor(full_treatment, level=x_order), y=di, fill=factor(group, level=measurement_order))) +
  geom_boxplot(alpha=0.5, outlier.size=0) +
  geom_point(aes(color=factor(group, level=measurement_order)), position = pj, size=1, show.legend = FALSE)+
  scale_fill_manual(values = c("#EBB134", "#ED6B5F", "#6060A8"), labels=c('OFAV Shallow', 'OFAV Deep', 'OFRA Deep')) + 
  scale_color_manual(values = c("#EBB134", "#ED6B5F", "#6060A8")) + 
  labs(y= "Coralite diameter (cm)", x = "Depth Treatment", fill='Origin') + 
  scale_x_discrete(labels=c('Deep', 'Shallow', 'Deep', 'Shallow', 'Deep', 'Shallow')) + 
  theme_classic() +
  stat_compare_means(comparisons = treatment_comparisons, method = "wilcox.test", 
                     symnum.args = list(cutpoints = c(0, 0.001 ,0.01, 0.05, Inf), symbols = c("***", "**", "*", "")))
Cdiameter <- ggplot(raw, aes(x=factor(full_treatment, level=x_order), y=Cdi, fill=factor(group, level=measurement_order))) +
  geom_boxplot(alpha=0.5, outlier.size=0) +
  geom_point(aes(color=factor(group, level=measurement_order)), position = pj, size=1, show.legend = FALSE)+
  scale_fill_manual(values = c("#EBB134", "#ED6B5F", "#6060A8"), labels=c('OFAV Shallow', 'OFAV Deep', 'OFRA Deep')) + 
  scale_color_manual(values = c("#EBB134", "#ED6B5F", "#6060A8")) + 
  labs(y= "Calice diameter (cm)", x = "Depth Treatment", fill='Origin') + 
  scale_x_discrete(labels=c('Deep', 'Shallow', 'Deep', 'Shallow', 'Deep', 'Shallow')) + 
  theme_classic() +
  stat_compare_means(comparisons = treatment_comparisons, method = "wilcox.test", 
                     symnum.args = list(cutpoints = c(0, 0.001 ,0.01, 0.05, Inf), symbols = c("***", "**", "*", "")))
diameter
Cdiameter
pairwise.wilcox.test(raw$di, raw$full_treatment, p.adjust.method="none")
pairwise.wilcox.test(raw$Cdi, raw$full_treatment, p.adjust.method="none")
```

### Results summary 

These results are EXACTLY the same as the area results, see that info. 


